Beyond Ultra-Diffuse Galaxy: Mass-Size Outliers in the Satellites of Milky Way Analogs

### 2022-09-08 main.tex
very ugly summary

\section{Summary}\label{sec:summary}
In this work, we present a comprehensive study for mass-size outliers including ultra-diffuse galaxies and ultra-puffy galaxies around Milky Way analogs at $0.01 < z < 0.04$.

We summarize our main findings and prospects below.

\begin{itemize}
    \item Using the HSC PDR2 data ($\sim 300\ \mathrm{deg}^{2}$), we conduct a systematic search for LSBGs using the method in \citet{Greco2018} complemented by our new deblending and modeling methods. Utilizing the non-parametric models generated by \code{scarlet}, we design a metric based on morphological parameters to effectively rule out false positives in the LSBG detection. We measure the structural properties of LSBGs using Spergel profiles and carefully characterize the measurement biases and uncertainties. The completeness of the search is also derived by injecting mock galaxies. Our LSBG search achieves high completeness among other searches and demonstrate the power of HSC data on LSB sciences.

    \item By matching LSBGs with MW analogs in NSA catalog, we construct samples of mass-size outliers

    \item their abundance

    \item their size and radial distribution

    \item their quenched fraction

\end{itemize}

we advocate studies for the quenched fraction as a function of size in simulations

Future works: nucleation, intrinsic shape, also the full LSBG sample of Johnny, LSBG cross-correlation, unsupervised machine learning, color gradient, non-par shapes.

>>>>

\section{Summary}\label{sec:summary}
In this work, we present a comprehensive study for mass-size outliers around Milky Way analogs at $0.01 < z < 0.04$. Beside the ultra-diffuse galaxies, we propose studying ``ultra-puffy galaxies'' which is defined to be $1.5\sigma$ above the average mass-size relation. We conduct statistical analysis for mass-size outliers and argue that UPGs better represent the large-size tail of the dwarf galaxy population. Surprisingly, we find that UPGs have similar quenched fraction as the bulk of MW satellites. We summarize our main findings and prospects below.

\begin{enumerate}

% \begin{itemize}
    \item Using the HSC PDR2 data ($\sim 300\ \mathrm{deg}^{2}$), we conduct a systematic search for LSBGs using the method in \citet{Greco2018} complemented by our new deblending and modeling methods. Utilizing the non-parametric models generated by \code{scarlet}, we design a metric based on morphological parameters to effectively rule out false positives in the LSBG detection. We measure the structural properties of LSBGs using Spergel profiles and carefully characterize the measurement biases and uncertainties. The completeness of the search is also derived by injecting mock galaxies. Our LSBG search achieves high completeness among other searches and demonstrate the power of HSC data on LSB sciences.

    \item By matching LSBGs with MW analogs in NSA catalog, we construct samples of mass-size outliers including UDGs and UPGs. We display their distributions on the size-surface brightness plane (Figure \ref{fig:udg_upg_re_mu}) and the mass-size plane (Figure \ref{fig:mass_size}). The UPG sample contains more bluer and higher surface brightness galaxies than the UDG sample. Instead, the UDG sample includes many red galaxies below $1.5\sigma$ from the average mass-size relation.

    \item After correcting for background contamination, the UDG abundance in MW analogs is $N_{\rm UDG} = 0.76\pm 0.04$ per host and the UPG abundance is $N_{\rm UPG} = 0.61 \pm 0.04$ per host. Our UDG abundance is slightly higher than that in ELVES, but much higher than that in SAGA. Combining data from studies in denser environments, we find that the UDG abundance follows a sublinear power law with the host halo mass. 
    \item We find the size distribution of UDGs follows a power law $n\propto r_e^{-1.60\pm0.21}$ which is shallower than the UDG size distribution in larger groups and clusters \citep{vdBurg2016,vdBurg2017}. We do not find significant dependence of size distribution on host stellar mass or the distance to host. The UPG size follows a unimodal distribution peaks at $\sim 2.5$ kpc.

    \item The radial distributions of UDGs and UPGs from their hosts can be described by projected NFW and Einasto profiles. We find that the best-fit NFW profiles are consistent with the NFW profile of a MW-mass halo, whereas the best-fit Einasto profiles have smaller concentrations than MW halos. The UDG and UPG follow similar radial distributions.

    \item We find that UDGs have high quenched fractions ($\sim 70\%$) and remain constant over 1.5 dex in satellite stellar mass (Figure \ref{fig:qfrac}). For UPGs, the quenched fraction is high at the lower-mass end $M_\star = 10^{6.8}\ M_\odot$, but the quenched fraction drops to $\sim 20\%$ at $M_\star = 10^{8.5}\ M_\odot$. Almost half of UPGs are quenched at $M_\star \approx 10^{8}\ M_\odot$. The quenched fractions of both UDGs and UPGs increase as satellites get closer to the hosts. Redder hosts also harbor more quenched UDGs and UPGs.

    \item Surprisingly, the quenched fraction of UPGs is very similar to the quenched fraction of the ``normal'' satellites in ELVES, although UPGs are mass-size outliers (Figure \ref{fig:qfrac_upg_elves}). We select ELVES UPGs and find they also have similar quenched fractions. The UPG quenched fraction also agrees with simulations (e.g., ARTEMIS \citealt{Font2022}, FIRE-2 \citealt{Samuel2022}) focusing on the bulk of MW satellites. The null trend of quenched fraction on the size of satellites indicates that quenching and size evolution of mass-size outliers happen separately in MW-like environments. It is plausible that these mass-size outliers are puffed up in the field before accretion, and quenched together with other MW satellites.

    \item We argue that the high and constant quenched fraction for UDGs is merely an artifact of the UDG definition which does not consider the color-$M_\star/L$ relation. We advocate studying UPG which is more physically motivated, does not introduce artifacts in the mass-size distribution, and better represents the large-size tail of dwarf galaxy population. 
% \end{itemize}
\end{enumerate}

This study on mass-size outliers only explores a small subset of LSBG candidates detected in HSC PDR2. In future works, we would like to exploit the full sample to study interesting topics including redshift distributions, nucleation fractions, intrinsic shapes of LSBGs. We will also explore machine learning techniques to help us classify LSBGs and estimate structural parameters. With the upcoming Vera Rubin Observatory Legacy Survey of Space and Time \citep[LSST, ][]{lsst2009,LSST2019}, we will be able to study the LSBG population with much greater statistics and gain new insights on the formation and evolution of mass-size outliers.
